\begin{problem}
  As we discussed in class, we can derive various ``cornerness'' metrics
  that take the form of functionals of \emph{only} the product
  and sum of the eigenvalues of the covariance matrix.
  
  \step
  Pick your favorite one (or propose your own),
  and explain how you would compute this metric efficiently
  for the entire image, using only convolutions and
  element-wise operations between images
  without explicitly computing eigenvalues.
  You can explain this either verbally, or using pseudocode.
\end{problem}

  % \bigskip
  
  % \step
\begin{answer}
  I chose to implement the Harris corner metric, which is defined as:
  \begin{align}
    \zaff{R(x,y)} &= \det \calM_w(x,y) - \kappa \cdot {(\tr \calM_w(x,y))}^2
  \end{align}
  where $\kappa$ is a constant.

  The main idea is that the product of a matrix's eigenvalues is equal to its determinant,
  and the sum of a matrix's eigenvalues is equal to its trace
  (the sum of elements along its diagonal).

  \begin{algorithm}[H]
    \caption{Compute the Harris Corner Metric for an Image}\label{alg:harris-metric}
    \begin{algorithmic}
      \Procedure{Harris}{$I,\;w,\; \kappa$}
        % sobel derivative filters
        \State
        \State $G_\sigma^x \gets \begin{bmatrix} 1 & 0 & -1 \\ 2 & 0 & -2 \\ 1 & 0 & -1 \end{bmatrix}$ \Comment{Sobel derivative filter.}
        \State
        \State $I_x \gets G_\sigma^x \ast I$
        \State $I_y \gets {(G_\sigma^x)}^T \ast I$
        % subtract mean from I_x and I_y
        \State $I_x \gets I_x - \mu(I_x)$ \Comment{Subtract mean from $I_x$.}
        \State $I_y \gets I_y - \mu(I_y)$ \Comment{Subtract mean from $I_y$.}
        \State $I_{x^2} \gets I_x \times I_x$
        \State $I_{y^2} \gets I_y \times I_y$
        \State $I_{xy} \gets I_x \times I_y$
        \State $R \gets [[0\; \mathbf{for}\;  y \in I]\; \mathbf{for}\; x \in I]$ \Comment{Initialize corner metric to zero for each pixel.}
        \For{$x \in I$}
          \For{$y \in I$}
            \State
            \State $M \gets w(x, y) \ast \begin{bmatrix} I_{x^2}[x,y] & I_{xy}[x,y] \\ I_{xy}[x,y] & I_{y^2}[x,y] \end{bmatrix}$ \Comment{Covariance matrix for current pixel.}
            \State
            \State $R[x,y] \gets \det M - \kappa \times {(\tr M)}^2$ \Comment{Corner metric for current pixel.}
          \EndFor
        \EndFor
        \For {$x \in I$}
          \For {$y \in I$}
            \For{$(i, j) \in \set{-1,0,1} \times \set{-1, 0, 1}$}
              \If{$R[x+i,y+j] > R[x,y]$}
                \State $R[x,y] \gets 0$ \Comment{Non-maximum suppression.}
                % break the inner loop.
                \State \textbf{break}
              \EndIf
            \EndFor
          \EndFor
        \EndFor
        \State \Return $R$
      \EndProcedure
    \end{algorithmic}
  \end{algorithm}
\end{answer}
