\begin{problem}
  \begin{enumroman}
    \item Prove that there exists a homography $\bH$ that satisfies
      \begin{equation}~\label{eq:1}
        \bx_1 \equiv \bH \bx_2
      \end{equation}
      between the 2D points (in homogeneous coordinates) $\bx_1$ and $\bx_2$
      in the images of a \emph{plane} $\Pi$ captured by two $3 \times 4$ camera
      projection matrices $\bP_1$ and $\bP_2$ respectively.
      The $\equiv$ symbol is equality up to scale.
      \emph{
        Note: A degenerate case happens when the plane $\Pi$ contains both cameras'
        centers, in which case there are infinite choices of $\bH$
        satisfying the above equation.
        You can ignore this special case in your answer.
      }

      \begin{answer}

      \end{answer}

    \item Prove that there exists a homography $\bH$ that satisfies equaton~\ref{eq:1}
      given two cameras separated by a pure rotation.
      That is, for camera $1$, $\bx_1 = \bK_1 \bmat{\bI & \bzero} \bX$,
      and for camera $2$, $\bx_2 = \bK_2 \bmat{\bR & \bzero} \bX$.
      Note that $\bK_1$ and $\bK_2$ are the $3 \times 3$ intrinsic matrices
      of the two cameras and are different. $\bI$ is the $3 \times 3$ identity matrix,
      $\bzero$ is the $3 \times 1$ zero vector, and $\bX$ is a point in 3D space.
      $\bR$ is the $3 \times 3$ rotation matrix of the camera.

      \begin{answer}

      \end{answer}

    \item Suppose that a camera is rotating about its center $\bC$,
      keeping the intrinsic parameters $\bK$ constant.
      Let $\bH$ be the homography that maps the view from one camera orientation
      to the view at a second orientation.
      Let $\theta$ be the angle of rotation between the two orientations.
      Show that $\bH^2$ is the homography corresponding to a rotation of $2\theta$.

      \begin{answer}

      \end{answer}
  \end{enumroman}
\end{problem}
