\begin{problem}
  In class, we say that a camera matrix satisfies the equation
  $\bx_i = \bP \bX_i$, and that six 3D-2D matches $\bx \leftrightarrow \bX$
  are sufficient to recover $\bP$ using a linear (non-iterative) algorithm.

  \step
  Find a linear algorithm for computing the camera matrix $\bP$
  in the special case when the camera location (but not orientation)
  is known. Ignoring degenerate configurations, how many 2D-3D matches
  are required for there to be a unique solution?
  Justify your answer.
\end{problem}

\begin{answer}
  In the special case when the camera location (but not orientation) is known,
  we can write the camera matrix $\bP$ as:
  \[ \bP = \bK \bmat{\bR & \bt} \]
  where $\bK$ is the $3 \times 3$ intrinsic matrix of the camera,
  $\bR$ is the $3 \times 3$ rotation matrix, and $\bt$ is the $3 \times 1$ translation vector,
  which is known.

  Our goal is to find $\bK$ and $\bR$ from the 3D-2D matches $\bx \leftrightarrow \bX$.
  Suppose we have $N$ 3D-2D matches $\bx \leftrightarrow \bX$.
  For any one correspondence $\bx_i \leftrightarrow \bX_i$, we can write:
  \begin{align*}
    \bx_i &= \bK \bmat{\bR &\bt} \bX_i \\
    \bmat{x_i \\ y_i \\ z_i} &= \bmat{f_x & 0 & c_x \\ 0 & f_y & c_y \\ 0 & 0 & 1}
    \bmat{
      r_1 & r_2 & r_3 & t_1 \\
      r_4 & r_5 & r_6 & t_2 \\
      r_7 & r_8 & r_9 & 1}
      \bmat{ X_i \\ Y_i \\ Z_i \\ 1} \\
    \intertext{So, we have:}
    &= \bmat{
      f_x r_1 + c_x r_7 & f_x r_2 + c_x r_8 & f_x r_3 + c_x r_9 & f_x t_1 + c_x \\
      f_y r_4 + c_y r_7 & f_y r_5 + c_y r_8 & f_y r_6 + c_y r_9 & f_y t_2 + c_y \\
      r_7 & r_8 & r_9 & 1} \bmat{X_i \\ Y_i \\ Z_i \\ 1} \\
  \end{align*}

  Thus, we get the following equations:
  \begin{align*}
    \bx_i &= (f_x r_1 + c_x r_7)\bX_i + (f_x r_2 + c_x r_8)\bY_i + (f_x r_3 + c_x r_9) \bZ_i + (f_x t_1 + c_x) \\
    \by_i &= (f_y r_4 + c_y r_7)\bX_i + (f_y r_5 + c_y r_8)\bY_i + (f_y r_6 + c_y r_9) \bZ_i + (f_y t_2 + c_y) \\
    \bz_i &= r_7 \bX_i + r_8 \bY_i + r_9 \bZ_i + 1
  \end{align*}

  $z_i$ is the scale factor, since
  \[ \bmat{\bx_i \\ \by_i \\ \bz_i} \equiv \bmat{\bx_i/\bz_i \\ \by_i/\bz_i \\ 1}. \]
  Thus, each 3D-2D match gives us two useful equations.
  Ignoring degenerate configurations where $\bK$ changes,
  and ignoring the scale factor, we would have $12 - 1 = 11$ unknowns in the matrix
  $\bmat{ \bR & \bt }$ if $\bt$ was unknown.
  However, since $\bt \equiv \bmat{t_1 \\ t_2 \\ 1}$ is known,
  we have $11-2 = 9$ remaining unknowns,
  so we need \emph{at least} $9$ equations to solve for the unknowns uniquely.
  Since each $\bx_i \leftrightarrow \bX_i$ correspondence gives us $2$ equations,
  we need at least $\ceil{\frac{9}{2}} = 5$ 3D-2D matches to have a unique solution.

  \step
  \textbf{Algorithm}
  \begin{enumarabic}
    \item For each 3D-2D match $\bx_i \leftrightarrow \bX_i$, write the equations
    \begin{align*}
      \bx_i &= (f_x r_1 + c_x r_7)\bX_i + (f_x r_2 + c_x r_8)\bY_i + (f_x r_3 + c_x r_9) \bZ_i + (f_x t_1 + c_x) \\
      \by_i &= (f_y r_4 + c_y r_7)\bX_i + (f_y r_5 + c_y r_8)\bY_i + (f_y r_6 + c_y r_9) \bZ_i + (f_y t_2 + c_y) \\
      \bz_i &= r_7 \bX_i + r_8 \bY_i + r_9 \bZ_i + 1
    \end{align*}
    \item Stack the equations for all $N = 5$ 3D-2D matches to get a system of linear equations
      \[ \bA \bR = \bb \]
    \item Solve for $\bR$ using the linear least squares method.
    \item Once $\bR$ is known, solve for $\bP$ using the equation
      \[ \bP = \bK \bmat{\bR & \bt} \]
  \end{enumarabic}
\end{answer}
